### 2018-11-07 algorithm.tex
fix #30: expectations over productivity shocks

\documentclass[12pt]{article}
\usepackage{amsmath}
\begin{document}
\section{Model solution}
Introduce a new variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\varrho_{njt} = \xi B_j T_{nj}^{-1/\theta} A_{njt}^{-1} w_{njt}^{\beta_j} \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}
With this variable, we can write the price of the final good as
\begin{equation}\label{eq:price}
	P_{mjt} = \left[
		\sum_{n=1}^N (\varrho_{njt}/\kappa_{mnjt})^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
This is a standard CES price index. We can write import shares as
\begin{equation}\label{eq:import_share}
	d_{mnjt} = \left[
	\frac
	{\varrho_{njt}/\kappa_{mnjt}}
	{P_{mjt}}
	\right]^{-\theta}.
\end{equation}
Wages can be expressed by inverting \eqref{eq:varrho},
\begin{equation}\label{eq:wage}
	w_{njt} = (\xi B_j)^{-1/\beta_j}
	\varrho_{njt}^{1/\beta_j}  
	T_{nj}^{1/(\beta_j\theta)}A_{njt}^{1/\beta_j} 
	\prod_{k=1}^J P_{nkt}^{-\gamma_{kj}/\beta_j}.
\end{equation}

@@

\subsection{Flexible labor equilibrium}
To get a starting point or direction of iteration, it is useful to calculate the flexible labor equilibrium, where wages are equal across sectors.

When wages are constant across sectors, we can write expenditure shares as
\begin{equation}\label{eq:sectoral_share:2}
e_{mjt} = \frac
	{w_{mt}\sum_{k=1}^J(\alpha_{mjt}(\mathbf P_t)\beta_k+\gamma_{jk}) L_{mkt}/\beta_k - S_{mt}}
	{w_{mt}\sum_{k=1}^J L_{mkt}/\beta_k-S_{mt}}.
\end{equation}
If $S_{mt}=0$ and demand is Cobb-Douglas, this can be precalculated in the outer loop, without knowing anything about wages. The inner loop (without the need for a middle loop) then solves for prices and wages.
\[
	\varrho_{njt} = w_{nt}^{-1/\theta}(L_{njt}/\beta_j)^{-1/\theta}
	\left[
	\sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		w_{mt}\sum_{k=1}^J(\alpha_{mjt}(\mathbf P_t)+\gamma_{jk}/\beta_k) L_{mkt} - S_{mt}
	\right]^{1/\theta}.
\]
We also know that 
\begin{equation}\label{eq:varrho:2}
	\varrho_{njt} = w_{nt}^{\beta_j} \xi B_j A_{njt}^{-1}  \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}

@@

\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\begin{document}
\subsection*{Normalizations and changes in notation}
I set $L_{nt}=1$ for all country and time. This is just a normalization, as the level of producitivity can pick up the slack.

I drop $T_n$. There is no need to carry a time invariant productivity around.

I use $A_{njt}$ rather than $Z_{njt}$. Because $L_{nt}=1$, the difference is only the exponent.

We used to have $y_{njt}/\psi_{njt}$ in the equation for productivity. This is only needed to express wages, so I just write out $w_{njt}$ explicitly in the formulas below.
\section{Model solution}
Introduce a new variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\varrho_{njt} = \xi B_j T_{nj}^{-1/\theta} A_{njt}^{-1} w_{njt}^{\beta_j} \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}
With this variable, we can write the price of the final good as
\begin{equation}\label{eq:price}
	P_{mjt} = \left[
		\sum_{n=1}^N (\varrho_{njt}/\kappa_{mnjt})^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
This is a standard CES price index. We can write import shares as
\begin{equation}\label{eq:import_share}
	d_{mnjt} = \left[
	\frac
	{\varrho_{njt}/\kappa_{mnjt}}
	{P_{mjt}}
	\right]^{-\theta}.
\end{equation}
Wages can be expressed by inverting \eqref{eq:varrho},
\begin{equation}\label{eq:wage}
	w_{njt} = (\xi B_j)^{-1/\beta_j}
	\varrho_{njt}^{1/\beta_j}  
	T_{nj}^{1/(\beta_j\theta)}A_{njt}^{1/\beta_j} 
	\prod_{k=1}^J P_{nkt}^{-\gamma_{kj}/\beta_j}.
\end{equation}

@@

\subsection{Calculating expectations}
Agents have rational expectations. The DGP of producitivities is as follows
\begin{equation}
	\ln A_{njt} = a_{njt}+\lambda_{jt} + \mu_{nt} +  \varepsilon_{njt}.
\end{equation}
The first term $a_{njt}$ is a deterministic trend, perfectly foreseen by agents. The second term $\lambda_{jt}$ is a global sectoral shock as in Koren and Tenreyro (2007). The third term $\mu_{nt}$ is a country shock, and $\varepsilon_{njt}$ is the idiosyncratic shock.

These shocks have the following DGP
\begin{align}
 	\lambda_{jt} &= \rho_{j}\lambda_{jt-1}+\sigma_{j}u_{jt}\\
 	\mu_{nt} &= \rho_{n}\mu_{nt-1}+\sigma_{n}u_{nt}\\
 	\varepsilon_{njt} &= \rho_{nj}\varepsilon_{njt-1}+\sigma_{nj}u_{njt}
\end{align} 
where the $u$s are serially and cross-sectionally independent standard normal variables. That is, each shock component follows an independent AR(1) process.

Expectations are taken conditional on $(a_{njt}, \lambda_{jt-1}, \mu_{nt-1})$, that is, over realizations of $(u_{jt}, u_{nt}, u_{njt})$. In practice we draw $S=100$ realizations of the innovations to compute the expected wage share.

Numerically integrating the expected wage shares over $u$s converges fast, and hence relatively low $S$ is sufficient, because wage shares are bounded between 0 and 1

@@

\subsection{Decomposing productivity shocks}
Given a series of productivities calibrated to the data, we decompose productivity into these four components:
\begin{equation}
	\ln A_{njt} = a_{njt}+\lambda_{jt} + \mu_{nt} +  \varepsilon_{njt}.
\end{equation}
The first term is a Baxter-King trend. The second is a global sector shock, calculated as the $(j,t)$ average of productivity deviations across countries. The third term is calculated as a weighted average of remaining shocks within a country-year. (Weighting is necessary because otherwise idiosyncratic shock of services, the larges sector, would dominate fluctuations in all countries.) The last term is a residual.

The parameter vector $\rho$ and $\sigma$ is estimate using simple OLS. Shocks $u$ are drawn for iid standard normal distribution.
\end{document}
